---
title: "Day 6: Stats Theory"
author: "Vasantha M"
---

## Introduction

> #### Overview
>
> -   t.test
>
> -   intuitive
>
> <!-- -->
>
> -   `set.seed() :` random number generation, good for reproducability - the number is a label for the set of numbers generated.
>
> -   `rnorm()` : generates random numbers from a normal (Guassian) distribution.
>
> -   `tibble()` : constructs a data frame.
>
> -   `gf_fitdistr()` : provides an easy way to overlay a probability distribution curve onto a plot: assessing how well the theoritical distribution matches the empirical data.

#### Setup Chunk:

```{r}
library(tidyverse)
library(mosaic)
library(ggformula)
library(infer)
library(broom) # Clean test results in tibble form
library(resampledata) # Datasets from Chihara and Hesterberg's book
library(openintro) # More datasets
```

# Inference for a Single Mean

.

## Case Study 1 - Toy Data

```{r}
set.seed(40)  # for replication
# Data as individual vectors ( for t.tests etc)
y <- rnorm(50, mean = 2, sd = 2)

# And as tibble too
mydata <- tibble(y = y)
mydata
```

.

## Inspecting and Charting Data:

.

```{r}
##
mydata %>%
    gf_density(~y) %>%
    gf_fitdistr(dist = "dnorm") %>%
    gf_labs(title = "Densities of Original Data Variables", subtitle = "Compared with Normal Density")
```

.

### Observations:

-   The variable y appear to be centered around 2

-   It does not seem to be normally distributed…

.

## Research Question

Could the mean of the population μ, from which y has been drawn, be zero?

## Assumptions

## Inference
