---
title: "(2) Day 2: Summaries"
author: "Vasantha M"
format: html
---

## Introduction:

How do we make data sets more readable? How can we decide how to group variables and compare them against each other in terms of what we are using the data for? This is my journey of exploring these questions with the help of [Data Summaries](https://av-quarto.netlify.app/content/courses/analytics/descriptive/modules/10-favstats/#how-do-these-summaries-work).

> ### Overview:
>
> -   Install and call packages to the document
>
> -   Select and call chosen data set
>
> -   Examining Data: Glimpse, Inspect, Skim and their different syntax
>
> -   Data Munging: mutate(), as_factor(), group_by() and summarise()
>
> -   Exploring babynames data set
>
> -   Exploring Startrek data set

.

.

### Set-up Chunk:

```{r}
#| label: setup
library(tidyverse)
library(ggformula)
library(mosaic)
library(skimr)
library(knitr)
library(kableExtra)
library(babynames)
```

.

.

> ### Notes
>
> -   Installing packages makes them available to all documents created in the future. {syntax: install.packages("")}
>
> -   In each document however, one must call on the packages using {library()} to "activate" them.
>
> -   Don't forget to add { #\| label: setup } in the first code chunk when calling the packages
>
> -   Have only one quarto document open in one window, as R treats a session as a single data pool and can pull libraries across documents (which could complicate things later.

.

.

## Looking at the 'mpg' Data Set:

```{r}
mpg
```

.

```{r}
mpg %>%
  head(10) %>%
  kbl(
    # add Human Readable column names
    col.names = c(
      "Manufacturer", "Model", "Engine\nDisplacement",
      "Model\n Year", "Cylinders", "Transmission",
      "Drivetrain", "City\n Mileage", "Highway\n Mileage",
      "Fuel", "Class\nOf\nVehicle"
    ),
    caption = "MPG Dataset"
  ) %>%
  kable_styling(
    bootstrap_options = c(
      "striped", "hover",
      "condensed", "responsive"
    ),
    full_width = F, position = "center"
  )
```

.

.

> ### Notes
>
> -   kbl - from the kableExtra package
>
> -   makes the data more human readable
>
> -   head(10) indicates the table will show 10 rows.
>
> -   '\\n' acts as line break
>
> -   **Styling (kable_styling):**
>
>     -   The table is styled using kable_styling with several bootstrap_options for appearance:
>
>         -   "striped": Alternating background colors on rows.
>
>         -   "hover": Highlights rows when hovering over them.
>
>         -   "condensed": Reduces the row height to make the table more compact.
>
>         -   "responsive": Ensures the table adjusts to different screen sizes.
>
>     -   full_width = F ensures that the table is not stretched to fill the entire width of the page but rather centered, as specified by position = "center".

.

.

### Using Glimpse

```{r}
glimpse(mpg)
```

#### *Testing different syntax:*

```{r}
mpg %>% dplyr::glimpse()
```

.

> ### Notes
>
> -   syntax essentially boils down to personal preference and readability of code.

.

### Using Inspect

```{r}
inspect(mpg)
```

.

### Using Skim

```{r}
skimr::skim(mpg)
```

#### *Testing different syntax:*

```{r}
skim(mpg)
```

.

.

## Data Munging - Mutating data according to your intent:

```{r}
mpg_modified <- mpg %>%
  dplyr::mutate(
    cyl = as_factor(cyl),
    fl = as_factor(fl),
    drv = as_factor(drv),
    class = as_factor(class),
    trans = as_factor(trans)
  )
glimpse(mpg_modified)
```

```{r}
inspect(mpg_modified)
```

.

> ### Notes
>
> -   Why make characters into factors? it better organizes the data in hand for plotting.
>
> -   Sometimes qual data such as year and date are considered as quant data due to their numeric content. These variables can be turned into qual data through munging.

.

### Grouping by Qualitative Variables

Highway milage grouped by cylinder type:

```{r}
mpg_modified %>%
  group_by(cyl) %>%
  summarize(average_hwy = mean(hwy), count = n())
```

The average milage of cars based on the cylinder type.

```{r}
mpg_modified %>%
  group_by(cyl, fl) %>%
  summarize(average_hwy = mean(hwy), count = n())
```

.

There should be 20 columns, but only there are only 8 as certain combinations of cars don't exisit in this data set.

.

.

## Looking at the 'babynames' Data Set:

```{r}
glimpse(babynames)
```

```{r}
## readable table for dataset babynames
babynames %>%
  head(20) %>%
  kbl(
    
    col.names = c(
      "Year", "Sex", "Name",
      "Count", "Proportion"
    ),
    caption = "babynames Dataset"
  ) %>%
  kable_styling(
    bootstrap_options = c(
      "striped", "hover",
      "condensed", "responsive"
    ),
    full_width = F, position = "center"
  )
```

```{r}
inspect(babynames)
```

```{r}
## transforming year from a quant variable to a qual variable
bn_modone <- babynames %>% 
  mutate(
    year = as_factor(year),
    name = as_factor(name)
  )
```

```{r}
## grouping by sex to see how many names fall under each category totally (playing around with androgenous names)

bn_modtwo <- bn_modone %>%
  group_by(sex) %>% 
  summarise(n_names = count(sex), Krishna = count(name == "Krishna"), Sam = count(name == "Sam"),Samuel = count(name == "Samuel"), Samantha = count(name == "Samantha"), Alex = count(name == "Alex"), Alexander = count(name == "Alexander"), Alexis = count(name == "Alexis"), Kai = count(name == "Kai"), Liza = count(name == "Liza"), Keethi_Keerti_kirthi_kirti = count(name == "Keerthi" | name == "Kirti"|name == "Keerti"|name == "Kirthi"), Manal = count(name == "Manal"))

glimpse(bn_modtwo)
```

```{r}
## trying to understand what exactly count(sex) means, as there are 'M' and 'F' both...
bn_modthree <- bn_modone %>%
  group_by(year) %>% 
  summarise(n_names = count(sex))

glimpse(bn_modthree)
```

.

> ### Notes
>
> -   What if I wanted to make 'year' into intervals of decades/centuries/half centuries and then make them into factors?
>
>     Arvind says "console\> ? case_when\> enter"
>
> -   What does the numerical output of count(sex) mean when grouped by year? Is it the count of only one sex? Or Both?

.

.

## Looking at Star Trek Data Set:

```{r}
#| eval: false
star_trek_data <- read_csv("../../Data/star_trek_books.csv")
star_trek_data
```

```{r}
star_trek_data <- read_delim(file = "../../Data/star_trek_books.csv", delim = ";")
star_trek_data
```

.

> ### Notes
>
> -   .csv files read commas as separators.
>
> -   When faced with any other kind of separators, use "read_delim", default for that is semicolon

.

```{r}
star_trek_data %>% glimpse()
```

### Making an easy-to-read Table:

```{r}
## removig the last coloumn for the sake of readability
startrek_data_mod <- star_trek_data %>%
  select(-dedication)

## formatting table for better readability
startrek_data_mod %>%
  head(20) %>%
  kbl(
    # add Human Readable column names
    col.names = c(
      "Title", "Author", "Date",
      "Publisher", "Identifier", "Series",
      "Subseries", "Number\n of\n Chapters", "Number\n of\n Words",
      "Number\n of\n Characters"
    ),
    caption = "Startrek Dataset"
  ) %>%
  kable_styling(
    bootstrap_options = c(
      "striped", "hover",
      "condensed", "responsive"
    ),
    full_width = F, position = "center"
  )
```

```{r}
star_trek_data %>% inspect()
```

```{r}
star_trek_data %>% skim()
```

.

.

## Munging Star Trek Data:

```{r}

## grouping by publisher to understand the average of certain quant variables
startrek_munged <- star_trek_data %>%
  group_by(publisher) %>%
  summarize(avg_nchap = mean(nchap), avg_nw = mean(nword), avg_nchar = mean(nchar), count = n())

## creating a readable table for this summary
startrek_munged %>%
  head(20) %>%
  kbl(
    col.names = c(
      "Publisher","Average\n #of\n Chapters", "Average\n #of\n Words", "Average\n #of\n Characters",
      "Total\n #of\n Books"
    ),
    caption = "Startrek Dataset - Munged"
  ) %>%
  column_spec(1 , border_right = TRUE) %>%
  kable_styling(
    bootstrap_options = c(
      "striped", "hover",
      "condensed", "responsive"
    ),
    full_width = F, position = "center"
  )
```

```{r}
##star_trek_data %>% 
  ##group_by(publisher) %>% 
  ##gitsummarize(author_count = sum(author))
```

.

.

## Conclusion:
