---
title: "(11) And Finally....."
author: "Vasantha M"
format: html
editor: visual
---

## Introduction

As I began writing what would be the last reflection for this class, a thought occurred to me. What better way to finally reflect, if not with graphs?

.

First I ranked each day of the course out of 10 on my `understanding` of the course material, my `learnings` and my `joy` levels. The following is a rundown of my code process to get to the graphs from this data, and then finally, my final final reflection :)

.

DISCLAIMER! - all inferences and data uncovered on this page must be taken with a mighty grain of salt. I've defined the rankings, and if anything, im a bit of a `va-riable` myself.

### Setup Chunk:

```{r}
#| label: setup
library(tidyverse)
library(kableExtra)
library(mosaic)
library(ggformula)
library(skimr)
library(ggbump)
library(ggplot2)
library(DT)
library(corrplot)
library(broom)
```

.

```{r}
## reading data
reflection <- read_csv(file ="../../Data/reflection.csv")
```

## Code

### Data:

```{r}
## data overview, as I initially input it
datatable(reflection, options = list(pageLength = 10))
```

.

```{r}
reflection %>% glimpse()
```

.

```{r}
reflection %>% inspect()
```

.

```{r}
reflection %>% skim()
```

.

.

### Graphs:

```{r}
## average of parameters across 10 days
reflection %>%
  drop_na() %>% 
  mutate( day = as_factor(day)) %>%
  summarize(learning = mean(learning),
            understanding = mean(understanding),
            joy = mean(joy)) %>%
   pivot_longer(cols = c("learning", "understanding", "joy"),
                    cols_vary = "slowest",
                    names_to = "parameter",
                    values_to ="meanvalue") %>%
  gf_col(meanvalue ~parameter,
         fill = ~parameter
  ) %>% 
  gf_refine(scale_fill_manual(values = c("learning" = "violet", "understanding" = "purple", "joy" = "pink"))) %>% 
  gf_labs(
    title = "Plot 1: Overall Parameters",
    caption = "25th October, 2024"
  )
  
  
  
```

.

#### Inference 1

Well, I seemed to have learned quite a lot, but haven't fully understood all I've learned. The joy in my learning is about as much as my understanding.

.

.

```{r}
## modifying data for stacked graphs

reflection_mod <- reflection %>%
  drop_na() %>% 
  mutate( day = as_factor(day)) %>% 
  pivot_longer(cols = c("learning", "understanding", "joy"),
                    cols_vary = "slowest",
                    names_to = "parameter",
                    values_to ="values") %>% glimpse()
  
  
```

.

```{r}
## Looking at the distribution of parameters

reflection_mod %>% 
  gf_col(day~values,
         fill = ~parameter) %>% 
   gf_refine(scale_fill_manual(values = c("learning" = "violet", "understanding" = "purple", "joy" = "pink"))) %>% 
  gf_labs(
    title = "Plot 2: Distribution of  Parameters",
    subtitle = "Stacked Graph",
    caption = "25th October, 2024"
  )
```

.

#### Inference 2

All three parameters gradually decreased as the days went on. Was this because I realized that we wouldn't be able to cover as much as I thought we would? Or was it because work picked up in terms of the cycle and I couldn't focus? Even I'm not entirely sure.

.

.

```{r}
## Looking at proportions for each day
reflection_mod %>% 
  gf_col(day~values,
         fill = ~parameter,
         position = "fill") %>% 
   gf_refine(scale_fill_manual(values = c("learning" = "violet", "understanding" = "purple", "joy" = "pink"))) %>% 
  gf_labs(
    title = "Plot 3: Proportion Distribution of Parameters",
    subtitle = "Filled Graph",
    caption = "25th October, 2024"
  )
```

.

#### Inference 3

The proportion of my joy, understanding and learning for each day seem to stay in the realm of each other, with nothing too extreme. Sure, there's definitely differences, but its safe to say that I expected and experienced the same of each day.

.

Interesting.

.

This means that even though I expected the same of each day, I didn't expect the same over 12 days, since I felt we couldn't cover as much. I guess this goes to say what a terrible long term judgement I have, oof :')

.

.

```{r}
## Looking at each parameter indivdually
reflection_mod %>% 
  gf_col(day~values,
         fill = ~parameter) %>% 
  gf_facet_wrap(~parameter) %>% 
   gf_refine(scale_fill_manual(values = c("learning" = "violet", "understanding" = "purple", "joy" = "pink"))) %>% 
  gf_labs(
    title = "Plot 4:  Individual Parameters by Day",
    subtitle = "Facet Wrapped",
    caption = "25th October, 2024"
  )
```

.

#### Inference 4

With the most gradual decline being joy and the most ups and downs in understanding, even if their averages are the same, its clear to see that learning was the most consistent (even if it followed similar trends) through the days. Nice!

.

.

```{r}
## Attempting a Bump Chart

ggplot(reflection_mod, aes(x = parameter, y = values, color = day, group = day)) +
  geom_bump(size = 1) +  
  geom_point(size = 4) +  
  scale_x_discrete(
    limits = c(  "learning", "understanding","joy"),
    labels = c( "learning", "understanding","joy" )
    ) + 
  labs(title = "Plot 5: Bump Chart Ranked by Parameter",
       caption = "25th October, 2024",
       x = "Parameters",
       y = "Rank",
       color = "") +
  scale_color_brewer(palette = "RdPu") +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.1))
        
```

.

#### Inference 5

The graph seems to have completely omitted rankings below 5, as all my rankings were 5 and above. In my experience, this class was above average, on the whole! Perhaps this isn't the best way to visualize this information, but it clearly sets apart the days with the most difference in ranking across parameter. Day 7, the first day of Stats Theory, must have really taken me by surprise.

.

.

```{r}
## How about the other way?
ggplot(reflection_mod, aes(x = day, y = values, color = parameter, group = parameter)) +
  geom_bump(size = 1) +  
  geom_point(size = 4) +  
  scale_x_discrete(
    limits = c("1", "2","3","4","5","7","8","9","10"),
    labels = c("D1", "D2","D3","D4","D5","D7","D8","D9","D10" )
    ) + 
  labs(title = "Plot 6: Bump Chart ranked by Day",
       caption = "25th October, 2024",
       x = "Day",
       y = "Ranks on 10",
       color = "Parameter") +
  scale_color_brewer(palette = "RdPu")+
  theme(axis.text.x = element_text(angle = 0, hjust = 0.1)) 
```

.

#### Inference 6

The other way in which a Bump Chart can be used for this data, it only confirms what we (I) have understood about myself.

.

.

```{r}
## correlation test for learning and understanding
mosaic::cor_test(learning ~ understanding, data = reflection) %>%
  broom::tidy() %>%
  knitr::kable(
    digits = 2,
    caption = "learning vs understanding"
  )
```

.

```{r}
## plotting correlation
reflection %>% 
  gf_point(learning ~ understanding) %>%
  gf_lm(color = "violet") %>%
  gf_labs(title = "Plot 7: Correlation between understanding and learning",
          x = "understanding",
          y = "learning")
```

.

#### Inference 7

As seen in the upwards slope of the graph, my learning and understanding have a positive correlation, with an estimate on 0.61. However, the confidence intervals include 0, indicating that it is not a significant correlation. We can't say for sure if my learning and understanding always depend on each other!

.

.

## Final (?) Reflection

This week really didn’t feel like the final week of class. Somehow, there is still so much more in the realm of R to be explored – I’m only just realizing that 12 days is barely anything in terms of the learning curve. Apart from R and RStudio, this class has taught me various things.

.

-   A blog for the web need not have to be some tediously hand-coded HTML document – the versatility of quarto takes care of all that.

    .

<!-- -->

-   I think I will also forever look at excel sheets differently; the potential in all those columns and rows for some seriously good art!

```         
.
```

-   Providing incentive goes a long way, acting as a catalyst for efficiency in data collection. Even better when the costs are shared, it’s definitely a good investment.

    .

-   Good design is informed by solid data. Hypothesis developed from the very same data is often built on pretences. That does not mean the conclusion isn’t credible: oftentimes, hypotheses are made up to be disproved, since doing the opposite isn’t feasible.

    .

Safe to say, this class has also left me with some questions:

.

-   Eliminating bias in terms of who the data is coming from is done with a coin toss. What about the data itself? How can a researcher account for false statements or partial truths?

    .

-   Trust – in both data and research process begs the question of ethics. Collaborative environments are inevitable, and navigating through them is sometimes more complicated than figuring out R. I can account for myself, but what about situations where I make up one of many parts? My group, my class, my institution?

    .

Collecting data ourselves was a lot more fun than I thought it would be. Generally, the thought of approaching random people acts as repellent enough from the task assigned, but just a few interviews in, the chocolates seemed to be performing their magic. It was also comforting to know that it did not matter who I approached – all was decided by the flip of a coin. The coin toss also became this sort of identifier: many guessed with an “Oh, are you in Arvind’s class?” before the question was even asked.

.

This won’t be goodbye, as there’s the whole of A3 to power through and hopefully some online classes as well. Super excited about all the resources I now have, especially those to pore over maps!!
