---
title: "Day 5: Groups"
author: "Vasantha M"
format: html
editor: visual
---

## Introduction

Today in class we discuss groups and box-plots.

> ### Overview:
>
> -   install new package 'palmerpenguins'

### Setup Chunk

```{r}
#| label: setup
library(tidyverse)
library(mosaic)
library(ggformula)
library(palmerpenguins)
```

## Examining Data:

```{r}
## Looking at the gss wages dataset
wages <- read_csv("https://vincentarelbundock.github.io/Rdatasets/csv/stevedata/gss_wages.csv")

##
wages
```

```{r}
wages %>% glimpse()
```

```{r}
wages %>% inspect()
```

## Munging Data:

> -   we can remove missing data in this case as there is enough data leftover to work with.

```{r}
wages_clean <-
  wages %>%
  tidyr::drop_na(realrinc) # choose column or leave blank to choose all
```

```{r}
wages_clean <- wages_clean %>%
  mutate(
    occrecode = as_factor(occrecode),
    wrkstat = as_factor(wrkstat),
    gender = as_factor(gender),
    educcat = as_factor(educcat),
    maritalcat = as_factor(maritalcat),
    rownames = as_factor(rownames),
    year = as_factor(year)
    
  )
wages_clean
```

```{r}
wages_clean %>% inspect()
```

## Q1: What is the basic distribution of realrinc (the respondant's base income)?

```{r}
wages_clean %>%
  gf_boxplot(realrinc ~ "Income") %>% # Dummy X-axis "variable"
  gf_labs(
    title = "Plot 1A: Income has a skewed distribution",
    subtitle = "Many outliers on the high side"
  )
```

> -   This graph has been plotted on a logarithmic scale (1e+05 = 1 x10\^5, 2e+05 = 2x10\^5 and so on) as the data spans a wide range of values
> -   **FROM THIS GRAPH**
>     -   Income is a very skewed distribution, as might be expected.
>     -   Presence of many higher-side outliers is noted.

## Q2: Is realrinc affected by gender?

```{r}
wages_clean %>%
  gf_boxplot(gender ~ realrinc) %>%
  gf_labs(title = "Plot 2A: Income by Gender")
```

> converting scalenatural log to log base 10. why??

```{r}
wages_clean %>%
  gf_boxplot(gender ~ log10(realrinc)) %>%
  gf_labs(title = "Plot 2B: Log(Income) by Gender")


```

```{r}
wages_clean %>%
  gf_boxplot(gender ~ realrinc, fill = ~gender) %>%
  gf_refine(scale_x_log10()) %>%
  gf_labs(title = "Plot 2C: Income filled by Gender, log scale")


```

> -   **FROM THE GRAPHS:**
>
>     -   Even when split by gender, realincome presents a skewed set of distributions.
>
>     -   The IQR for males is smaller than the IQR for females. There is less variation in the middle ranges of realrinc for men.
>
>     -   log10 transformation helps to view and understand the regions of low realrinc.
>
>     -   There are outliers on both sides, indicating that there may be many people who make very small amounts of money and large amounts of money in both genders.
