---
title: "Day 3: Counts"
author: "Vasantha M"
format: html
editor: visual
---

## Introduction:

Today we as a class explored [Counts](https://av-quarto.netlify.app/content/courses/analytics/descriptive/modules/20-barplots/) using a data set, taxi, about taxi rides taken in Chicago in 2022. Under the 'Your Turn' section of the [Counts](https://av-quarto.netlify.app/content/courses/analytics/descriptive/modules/20-barplots/) page, I chose to investigate the dataset.

> ### Overview
>
> -   First install packages and call them to the document
>
> -   Then call the data set and perform glimpse(), inspect() and skim()
>
> -   If there is anything to be munged, the data is re-organised
>
> -   Looking at the data, questions are developed that can be answered by comparing variables
>
> -   Then, plot graphs, compare and analyze them
>
> -   Lastly revisit the entire process with a new dataset!

.

.

### Setup Chunk:

```{r}
#| label: setup
library(tidyverse)
library(mosaic)
library(ggformula)
library(skimr)
library(kableExtra)
```

.

.

## Reading Data:

```{r}
taxi <- read_csv("https://vincentarelbundock.github.io/Rdatasets/csv/modeldata/taxi.csv")
taxi
```

```{r}
taxi %>% glimpse()
```

```{r}
##TABLE!!
taxi %>%
  head(20) %>%
  kbl(
   
    col.names = c(
      "SN", "Tip", "Distance",
      "Company", "Local", "DOW",
      "Month", "Hour"
    ),
    caption = "Taxi Dataset"
  )%>%
  kable_styling(
    bootstrap_options = c(
      "striped", "hover",
      "condensed", "responsive"
    ),
    full_width = F, position = "center"
  )
```

```{r}
taxi %>%inspect()
```

```{r}
taxi %>% skim()
```

.

.

## Munging Data

.

### Converting Qualitative Variables into Factors:

```{r}
taxi_modifiedA <- taxi %>%
  dplyr::mutate(
    tip = as_factor(tip),
    local = as_factor(local),
    month = as_factor(month),
    dow = as_factor(dow),
    company = as_factor(company),
    hour = as_factor(hour)
    
  )
glimpse(taxi_modifiedA)
```

```{r}
taxi_modifiedA %>%inspect()
```

### Ordered Factors - Levels and Labels

```{r}

## Convert `dow`, `local`, and `month` into ordered factors
taxi_modifiedB <- taxi_modifiedA %>%
  mutate(
    dow = factor(dow,
      levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"),
      labels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"),
      ordered = TRUE
    ),
    ##
    local = factor(local,
      levels = c("yes", "no"),
      labels = c("yes", "no"),
      ordered = TRUE
    ),
    ##
    month = factor(month,
      levels = c("Jan", "Feb", "Mar", "Apr"),
      labels = c("Jan", "Feb", "Mar", "Apr"),
      ordered = TRUE
    )
  )
taxi_modifiedB %>% glimpse()
```

> ### Notes
>
> -   when ordered isn't set to true, the labels and levels resort to alphabetical as default

.

.

```{r}
taxi_modifiedB %>% inspect()
```

.

.

## Q1.A: Do more people tip than not?

```{r}
gf_bar(~tip, data = taxi_modifiedB) %>%
  gf_labs(title = "Plot 1: Counts of Tips")
```

.

.

> ### Notes
>
> -   color vs fill
>
> -   where are the colours coming from? How can I change them
>
> -   position: dodge vs stacked - DEFAULT IS STACKED
>
> -   gf_bar - when the y is left blank, it counts
>
> -   gf_coloumn - gives the y axis to manually input

.

.

## Q2.A: Does the tip depend upon whether the trip is local or not?

```{r}
## using colour = ...
taxi_modifiedB %>%
  gf_bar(~local,
    color = ~tip,
    position = "dodge"
  ) %>%
  gf_labs(title = "Plot 2A.1: Dodged Bar Chart",
          subtitle = "Using Colour")
```

.

```{r}
## Using fill = ...
taxi_modifiedB %>%
  gf_bar(~local,
    fill = ~tip,
    position = "dodge"
  ) %>%
  gf_labs(title = "Plot 2A.2: Dodged Bar Chart",
          subtitle = "Using Fill")
```

.

```{r}
## Flipping facets
taxi_modifiedB %>%
  gf_bar(~tip,
    fill = ~local,
    position = "dodge"
  ) %>%
  gf_labs(title = "Plot 2A.3: Dodged Bar Chart",
          subtitle = "Flipping 'local' and 'tip' variables:")
```

.

```{r}
## Using stack
taxi_modifiedB %>%
  gf_bar(~local,
    fill = ~tip,
    position = "stack"
  ) %>%
  gf_labs(
    title = "Plot 2B.1: Stacked Bar Chart",
    subtitle = "Can we spot per group differences in proportions??"
  )
```

.

```{r}
## Flipping Facets
taxi_modifiedB %>%
  gf_bar(~tip,
    fill = ~local,
    position = "stack"
  ) %>%
  gf_labs(
    title = "Plot 2B.2: Stacked Bar Chart",
    subtitle = "Flipping 'local' and 'tip' variables:"
  )

```

.

```{r}
## Using ratio proportions
taxi_modifiedB %>%
  gf_bar(~local,
    fill = ~tip,
    position = "fill"
  ) %>%
  gf_labs(
    title = "Plot 2C: Filled Bar Chart",
    subtitle = "Comparing using perecentages and ratios"
  )
```

.

.

> ### Understandings from Graphs:
>
> -   Plot 2B. 1: People tip more often than not, regardless of if the trip is local or not.
>
> -   Plot 2B. 2: Majority of those who tip are going on trips that are not local. This may be true, but its not the *whole truth*.
>
>     *BECAUSE PERCENTAGS MATTER - THE NUMBER OF TRIPS THAT ARE LOCAL AND NOT LOCAL ARE NOT THE SAME.*
>
> -   Plot 2C: Now we can say for sure that more people tip on trips that are not local.

.

.

## Q3: Do some cab companies get more tips than others?

```{r}
taxi_modifiedB %>%
  gf_bar(~ company,
    fill = ~tip,
    position = "fill"
  ) %>%
  gf_labs(
    title = "Plot 3A: Filled Bar Chart",
    subtitle = "Comparing using perecentages and ratios by company"
  )
```

.

### Re-Ordering the columns.

```{r}
taxi_modifiedC <- taxi_modifiedB %>% 
  group_by(company) %>%
  summarise(tipyes_prop = prop(tip), tipno_prop= 1-prop(tip), count_tip = count(tip))%>% arrange(desc(tipyes_prop))
taxi_modifiedC

```

```{r}
taxi_modifiedC$company <- factor(taxi_modifiedC$company,
                           levels = taxi_modifiedC$company[order(-taxi_modifiedC$tipyes_prop)])
  
taxi_modifiedC %>% glimpse()
```

```{r}
##a = taxi_modifiedC$tipyes_prop
##b = taxi_modifiedC$tipno_prop
##tips = a+b
##taxi_modifiedD <- taxi_modifiedC %>%
  ##dplyr::mutate(
    ##tip = factor(tip,
                 ##levels = c("yes","no"),
                 ##labels
  ##)
  
```

```{r}
taxi_modifiedC %>%
  gf_bar(~ company,
    fill = ~ tipyes_prop,
    position = "fill"
  ) %>% 
  gf_labs(
    title = "Plot 3A: Filled Bar Chart",
    subtitle = "Comparing using perecentages and ratios by company"
  ) %>% 
  
  gf_theme(theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    )
  ))

```

.

> ### Notes
>
> -   using arrange(), select() and reorder()
>
> -   Don't mix in BaseR when you have tidyverse!!
>
> -   How can I bring back the yes and no comparison of tips within each coloumn?

.

## Q4: **Does a tip depend upon the distance, hour of day, and DOW and month?**

.

### Tips by Hour

```{r}
taxi_modifiedB %>%
  gf_bar(~ hour,
    fill = ~tip,
    position = "fill"
  ) %>%
  gf_labs(
    title = "Plot 4A: [Tips and Hour] Filled Bar Chart",
    subtitle = ".."
  )
```

.

### Tips by DOW

```{r}
taxi_modifiedB %>%
  gf_bar(~ dow,
    fill = ~tip,
    position = "fill"
  ) %>%
  gf_labs(
    title = "Plot 4B: [Tips and DOW] Filled Bar Chart",
    subtitle = ".."
  )
```

.

### Tips by month

```{r}
taxi_modifiedB %>%
  gf_bar(~ month,
    fill = ~tip,
    position = "fill"
  ) %>%
  gf_labs(
    title = "Plot 4C: [Tips and Month] Filled Bar Chart",
    subtitle = ".."
  )
```

.

### Tips by DOW and Month

```{r}
## With position default
 taxi_modifiedB %>% 
   gf_bar(~ month | dow, 
       fill = ~tip) %>%
  gf_labs(
    title = "Plot E: Counts of Tips by Day of Week and Month")
```

.

```{r}
## With position fill
 taxi_modifiedB %>% 
   gf_bar(~ month | dow, 
       fill = ~tip,
       position = "fill") %>%
  gf_labs(
    title = "Plot E: Counts of Tips by Day of Week and Month",
    subtitle = "According to ratio and proportions")
```

.

### Tips by DOW an Hour

```{r}

taxi_modifiedB %>% 
  gf_bar(~ hour | dow, 
         fill = ~tip) %>%
  gf_labs(
    title = "Plot G: Counts of Tips by Hour and Day of Week",
    
  )

## swapping facets would result in a graph that is harder to grasp
```

.

.

## Conclusion:
