---
title: "Day 3: Counts"
author: "Vasantha M"
format: html
editor: visual
---

## Introduction:

Today we as a class explored [Counts](https://av-quarto.netlify.app/content/courses/analytics/descriptive/modules/20-barplots/) using a data set, taxi, about taxi rides taken in Chicago in 2022. Under the 'Your Turn' section of the [Counts](https://av-quarto.netlify.app/content/courses/analytics/descriptive/modules/20-barplots/) page, I chose to investigate the fertility data set.

> ### Overview
>
> -   First install packages and call them to the document
>
> -   Then call the data set and perform glimpse(), inspect() and skim()
>
> -   If there is anything to be munged, the data is re-organised
>
> -   Looking at the data, questions are developed that can be answered by comparing variables
>
> -   Then, plot graphs, compare and analyze them
>
> -   Lastly revisit the entire process with a new dataset!

.

.

### Setup Chunk:

```{r}
#| label: setup
library(tidyverse)
library(mosaic)
library(ggformula)
library(skimr)
library(kableExtra)
```

.

.

## Reading Data:

```{r}
taxi <- read_csv("https://vincentarelbundock.github.io/Rdatasets/csv/modeldata/taxi.csv")
taxi
```

```{r}
taxi %>% glimpse()
```

```{r}
##TABLE!!
taxi %>%
  head(20) %>%
  kbl(
   
    col.names = c(
      "SN", "Tip", "Distance",
      "Company", "Local", "DOW",
      "Month", "Hour"
    ),
    caption = "Taxi Dataset"
  )%>%
  kable_styling(
    bootstrap_options = c(
      "striped", "hover",
      "condensed", "responsive"
    ),
    full_width = F, position = "center"
  )
```

```{r}
taxi %>%inspect()
```

```{r}
taxi %>% skim()
```

.

.

## Munging Data

.

### Converting Qualitative Variables into Factors:

```{r}
taxi_modifiedA <- taxi %>%
  dplyr::mutate(
    tip = as_factor(tip),
    local = as_factor(local),
    month = as_factor(month),
    dow = as_factor(dow),
    company = as_factor(company),
    hour = as_factor(hour)
    
  )
glimpse(taxi_modifiedA)
```

```{r}
taxi_modifiedA %>%inspect()
```

### Ordered Factors - Levels and Labels

```{r}

## Convert `dow`, `local`, and `month` into ordered factors
taxi_modifiedB <- taxi_modifiedA %>%
  mutate(
    dow = factor(dow,
      levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"),
      labels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"),
      ordered = TRUE
    ),
    ##
    local = factor(local,
      levels = c("yes", "no"),
      labels = c("yes", "no"),
      ordered = TRUE
    ),
    ##
    month = factor(month,
      levels = c("Jan", "Feb", "Mar", "Apr"),
      labels = c("Jan", "Feb", "Mar", "Apr"),
      ordered = TRUE
    )
  )
taxi_modifiedB %>% glimpse()
```

> ### Notes
>
> -   when ordered isn't set to true, the labels and levels resort to alphabetical as default

.

.

```{r}
taxi_modifiedB %>% inspect()
```

.

.

## Q1.A: Do more people tip than not?

```{r}
gf_bar(~tip, data = taxi_modifiedB) %>%
  gf_labs(title = "Plot 1: Counts of Tips")
```

.

.

> ### Notes
>
> -   color vs fill
>
> -   where are the colours coming from? How can I change them
>
> -   position: dodge vs stacked - DEFAULT IS STACKED
>
> -   gf_bar - when the y is left blank, it counts
>
> -   gf_coloumn - gives the y axis to manually input

.

.

## Q2.A: Does the tip depend upon whether the trip is local or not?

```{r}
## using colour = ...
taxi_modifiedB %>%
  gf_bar(~local,
    color = ~tip,
    position = "dodge"
  ) %>%
  gf_labs(title = "Plot 2A.1: Dodged Bar Chart",
          subtitle = "Using Colour")
```

.

```{r}
## Using fill = ...
taxi_modifiedB %>%
  gf_bar(~local,
    fill = ~tip,
    position = "dodge"
  ) %>%
  gf_labs(title = "Plot 2A.2: Dodged Bar Chart",
          subtitle = "Using Fill")
```

.

```{r}
## Flipping facets
taxi_modifiedB %>%
  gf_bar(~tip,
    fill = ~local,
    position = "dodge"
  ) %>%
  gf_labs(title = "Plot 2A.3: Dodged Bar Chart",
          subtitle = "Flipping 'local' and 'tip' variables:")
```

.

```{r}
## Using stack
taxi_modifiedB %>%
  gf_bar(~local,
    fill = ~tip,
    position = "stack"
  ) %>%
  gf_labs(
    title = "Plot 2B.1: Stacked Bar Chart",
    subtitle = "Can we spot per group differences in proportions??"
  )
```

.

```{r}
## Flipping Facets
taxi_modifiedB %>%
  gf_bar(~tip,
    fill = ~local,
    position = "stack"
  ) %>%
  gf_labs(
    title = "Plot 2B.2: Stacked Bar Chart",
    subtitle = "Flipping 'local' and 'tip' variables:"
  )

```

.

```{r}
## Using ratio proportions
taxi_modifiedB %>%
  gf_bar(~local,
    fill = ~tip,
    position = "fill"
  ) %>%
  gf_labs(
    title = "Plot 2C: Filled Bar Chart",
    subtitle = "Comparing using perecentages and ratios"
  )
```

.

.

> ### Understandings from Graphs:
>
> -   Plot 2B. 1: People tip more often than not, regardless of if the trip is local or not.
>
> -   Plot 2B. 2: Majority of those who tip are going on trips that are not local. This may be true, but its not the *whole truth*.
>
>     *BECAUSE PERCENTAGS MATTER - THE NUMBER OF TRIPS THAT ARE LOCAL AND NOT LOCAL ARE NOT THE SAME.*
>
> -   Plot 2C: Now we can say for sure that more people tip on trips that are not local.

.

.

## Q3: Do some cab companies get more tips than others?

```{r}
taxi_modifiedB %>%
  gf_bar(~ company,
    fill = ~tip,
    position = "fill"
  ) %>%
  gf_labs(
    title = "Plot 3A: Filled Bar Chart",
    subtitle = "Comparing using perecentages and ratios by company"
  )
```

.

### Re-Ordering the columns.

```{r}
taxi_modifiedC <- taxi_modifiedB %>% 
  group_by(company) %>%
  summarise(tipyes_prop = prop(tip), tipno_prop= 1-prop(tip), count_tip = count(tip))%>% arrange(desc(tipyes_prop))
taxi_modifiedC

```

```{r}
taxi_modifiedC$company <- factor(taxi_modifiedC$company,
                           levels = taxi_modifiedC$company[order(-taxi_modifiedC$tipyes_prop)])
  
taxi_modifiedC %>% glimpse()
```

```{r}
##a = taxi_modifiedC$tipyes_prop
##b = taxi_modifiedC$tipno_prop
##tips = a+b
##taxi_modifiedD <- taxi_modifiedC %>%
  ##dplyr::mutate(
    ##tip = factor(tip,
                 ##levels = c("yes","no"),
                 ##labels
  ##)
  
```

```{r}
taxi_modifiedC %>%
  gf_bar(~ company,
    fill = ~ tipyes_prop,
    position = "fill"
  ) %>% 
  gf_labs(
    title = "Plot 3A: Filled Bar Chart",
    subtitle = "Comparing using perecentages and ratios by company"
  ) %>% 
  
  gf_theme(theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    )
  ))

```

.

> ### Notes
>
> -   using arrange(), select() and reorder()
>
> -   Don't mix in BaseR when you have tidyverse!!
>
> -   How can I bring back the yes and no comparison of tips within each coloumn?

.

## Q4: **Does a tip depend upon the distance, hour of day, and DOW and month?**

.

### Tips by Hour

```{r}
taxi_modifiedB %>%
  gf_bar(~ hour,
    fill = ~tip,
    position = "fill"
  ) %>%
  gf_labs(
    title = "Plot 4A: [Tips and Hour] Filled Bar Chart",
    subtitle = ".."
  )
```

.

### Tips by DOW

```{r}
taxi_modifiedB %>%
  gf_bar(~ dow,
    fill = ~tip,
    position = "fill"
  ) %>%
  gf_labs(
    title = "Plot 4B: [Tips and DOW] Filled Bar Chart",
    subtitle = ".."
  )
```

.

### Tips by month

```{r}
taxi_modifiedB %>%
  gf_bar(~ month,
    fill = ~tip,
    position = "fill"
  ) %>%
  gf_labs(
    title = "Plot 4C: [Tips and Month] Filled Bar Chart",
    subtitle = ".."
  )
```

.

### Tips by DOW and Month

```{r}
## With position default
 taxi_modifiedB %>% 
   gf_bar(~ month | dow, 
       fill = ~tip) %>%
  gf_labs(
    title = "Plot E: Counts of Tips by Day of Week and Month")
```

.

```{r}
## With position fill
 taxi_modifiedB %>% 
   gf_bar(~ month | dow, 
       fill = ~tip,
       position = "fill") %>%
  gf_labs(
    title = "Plot E: Counts of Tips by Day of Week and Month",
    subtitle = "According to ratio and proportions")
```

.

### Tips by DOW an Hour

```{r}

taxi_modifiedB %>% 
  gf_bar(~ hour | dow, 
         fill = ~tip) %>%
  gf_labs(
    title = "Plot G: Counts of Tips by Hour and Day of Week",
    
  )

## swapping facets would result in a graph that is harder to grasp
```

.

.

## 'Your Turn': Fertility Data Set:

.

### Examining Data:

```{r}
fertility <- read_csv( file = "../../Data/Fertility.csv")
fertility%>%glimpse()
```

```{r}
## Table
fertility %>%
  head(20)%>%
  kbl(
   
    col.names = c(
      "SN", "More\n Children", "First\n Child",
      "Second\n Child", "Age", "African\n American","Hispanic", "Other", "Work\n in Hours"
      
    ),
    caption = "Fertility Dataset"
  )%>%
  kable_styling(
    bootstrap_options = c(
      "striped", "hover",
      "condensed", "responsive"
    ),
    full_width = F, position = "center"
  )
```

```{r}
fertility %>% inspect()
```

.

.

### Data Dictionary:

I tried finding where the data set originated, but couldn't really trace anything back. From what I could make sense of it, these are my assumptions:

-   There `2,54,654` participants that make up this dataset, with no missing entries.

<!-- -->

-   `morekids`: Those who want more kids, either `yes` or `no`

-   `gender1`: Gender of the first child

-   `gender2`: Gender of the second child

-   `age`: age of the participant in years

-   `afam`: if the participant identifies as African American, either `yes` or `no`

-   `hispanic`: if the participant identifies as Hispanic, either `yes` or `no`

-   `other`: if the participant identifies as another minority, either `yes` or `no`

-   `work:` the number of hours worked in a week

.

.

### Munging Data:

```{r}
fertility_modified <- fertility%>%
  mutate(
    morekids = as_factor(morekids),
    gender1 = as_factor(gender1),
    gender2 = as_factor(gender2),
    afam = as_factor(afam),
    hispanic = as_factor(hispanic),
    other = as_factor(other),
  )
glimpse(fertility_modified)
```

```{r}
fertility_modified %>%
  dplyr::mutate(
    morekids = factor(morekids,
      levels = c("yes", "no"),
      labels = c("Yes", "No"),
      ordered = TRUE
    ),
    ##
    gender1 = factor(gender1,
      levels = c("male", "female"),
      labels = c("Male", "Female"),
      ordered = TRUE
    ),
    ##
    gender2 = factor(gender2,
      levels = c("male", "female"),
      labels = c("Male", "Female"),
      ordered = TRUE
    ),
    ##
    afam = factor(afam,
      levels = c("yes", "no"),
      labels = c("Yes", "No"),
      ordered = TRUE
    ),
     ##
    hispanic = factor(hispanic,
      levels = c("yes", "no"),
      labels = c("Yes", "No"),
      ordered = TRUE
    ),
     ##
    other = factor(other,
      levels = c("yes", "no"),
      labels = c("Yes", "No"),
      ordered = TRUE
    )
  )
fertility_modified %>% glimpse()
  
```

.

### Q0: How many participants want more kids?

```{r}
## Count of those who want more kids
fertility_modified %>% gf_bar(~morekids) %>%
  gf_labs(title = "Plot 0.A: Counts of those who want More Kids",
          subtitle = "fill = none")
```

.

```{r}
## Propotion of those who want more kids
fertility_modified %>% 
  gf_bar(~morekids,
         fill = ~morekids,
          position = "stack"           ) %>%
  gf_labs(title = "Plot 0.B: Proportion of those who want More Kids",
          subtitle = "position = 'fill'")
```

.

> Essentialy I wanted a graph with a single bar, but toggling with 'position' doesn't seem to be doing anything. Why?
>
> But moving on:

.

.

### Q1: How does the number of working hours affect the likelihood of wanting more kids?

.

```{r}
## trying to use gf_col and make the y axis working hours
fertility_modified %>%
  drop_na(work) %>% 
  gf_col(work ~ morekids) %>%
  gf_labs(title = "Plot 1.A: Those who want More Kids plotted against Working Hours",
          subtitle = "fill = none")
```

> not really the result I was looking for; why is the y axis in a logarithmic scale??

```{r}
## trying with gf_bar and fill instead
fertility_modified %>% 
  gf_bar(~work,
         fill = ~morekids,
         position = "fill") %>%
  gf_labs(title = "Plot 1.B: Those who want More Kids by Working Hours",
          subtitle = "fill = morekids")
```

.

> Now that's more like it!!
>
> How would this look like as a line graph?

.

```{r}
## work vs morekids line graph
fertility_modified %>%
  group_by(work) %>% 
  summarize(morekidsprop = mean(morekids == "yes")) %>% 
  gf_line(morekidsprop ~ work)%>%
  gf_labs(title = "Plot 1.C:  Proportion of Those who want More Kids by Working Hours",
          subtitle = "line graph")
  
```

.

> -   **Understandings from the Graphs:**
>
>     -   Plot 0.A: In general, there are more people who do not wan;t kids.
>
>     -   Plot 1.B: There seems to be a gradual decline in those who want kids as the work hours per week increase
>
>     -   Plot 1.C: The line graph indicates that there is indeed a decline overall, though it is not a smooth one. When looking at those who work 50 hours a week, the proportion of those who want kids is definitely less than those who work 0 hours a week - However, the same graph indicates that those who work 25 hours a week have a higher proportion of those who want kids in comparison to those who work 23 hours (a lesser amount).
>
> -   Lets plot other variables to see if there are any stronger patterns

.

.

### Q2: Does age affect the likelihood of wanting more kids?

.

```{r}
## plotting age vs morekids
fertility_modified %>% 
  gf_bar(~age,
         fill = ~morekids,
         position = "fill") %>%
  gf_labs(title = "Plot 2: Those who want More Kids by their age",
          subtitle = "fill = morekids")
```

.

> Wow, interestingly, the proportion of those who want more kids increases steadily with age.

.

.

### Q3: How do age and work factor into wanting more kids?

.

```{r}
## plotting morekids against both age and work
fertility_modified %>%
  gf_bar(~ work | age,
         fill = ~morekids,
         position = "fill") %>%
  gf_labs(title = "Plot 3.A: Those who want More Kids by Age and Working Hours",
          subtitle = "fill = morekids")
 
```

> The readability is quite low.. what if i filter the ages to look at individually?

```{r}
## plotting morekids against both age and work - filtered
fertility_modified %>% 
  filter(age == "21")%>%
  gf_bar(~ work | age,
         fill = ~morekids,
         position = "fill") %>%
  gf_labs(title = "Plot 3.B: Those who are 21 vs More Kids and Working Hours",
          subtitle = "fill = morekids")
 
```

.

> -   **Understandings from the Graphs:**
>
>     -   Plot 3.A: Those who are younger seem to have more variation in terms of wanting more kids, that is, it does not seem to be related to their hours of work. Those who are older seem to be of similar opinions across different amounts of work hours, but there is still no established pattern in the relation of the two variables
>
>     -   Plot 3.B: The missing bars indicate that there is no data at that intersection, i.e, there is nobody in this data set who is 21 years old and works 21 hours a week, and so on.

.

.

### Q4: Do the gender of the existing children play a part in wanting more kids?

.

```{r}
## plotting distribution of gender1 and gender 2
fertility_modified %>% 
  gf_bar(~ gender1 | gender2,
         fill = ~morekids,
         position = "stack") %>%
  gf_labs(title = "Plot 4.A: morekids vs gender",
          subtitle = "fill = morekids")
```

> How can this graph be made more readable?
>
> Lets try.

```{r}
fertility_modifiedA <- fertility_modified %>%
  dplyr::mutate(
    gender = paste( gender1, gender2, sep = "_"),
    gender = factor(gender,
                    levels = c("male_female", "male_male", "female_male", "female_female", ordered = TRUE ),
                   labels = c("male&female", "male&male","male&female", "female&female", ordered = TRUE )))
 

fertility_modifiedA

```

```{r}
fertility_modifiedA%>% 
  gf_bar(~morekids,
         fill = ~gender,
         position = "dodge"
         )%>%
  gf_labs(title = "Plot 4.B: morekids vs gender",
          subtitle = "fill = gender")
```

> Success!
>
> Now in looking at proportion:

```{r}
fertility_modifiedA%>% 
  gf_bar(~morekids,
         fill = ~gender,
         position = "fill"
         )%>%
  gf_labs(title = "Plot 4.C: morekids vs gender",
          subtitle = "fill = gender")
```

.

> -   **Understandings from the Graphs:**
>
>     -   Plot 4.A: The three combinations of gender are the same relative to each other in both 'yes' and 'no' sections under wanting more kids. in other words, both sections have participants with male and female children as the largest proportion and participants with two female kids making up the smallest proportion
>
>     -   Plot 4.B: Comparing the proportions, however, can tell us that **those who have two kids of the same gender already are more likely (even if by a small margin) to want more kids.**

.

.

### Q5. What relation (if any) is there between ethnicity and wanting to have more kids?

.

```{r}
fertility_modified %>% 
  gf_bar(~morekids,
         fill = ~afam,
                              )%>%  
gf_labs(
  title = "Plot 5.A: Counts of those who want More Kids",
  subtitle = "fill = afam")
```

.

> Calculating proportions:

.

```{r}
fertility_modified %>%
  ##
  group_by(morekids) %>% 
  ##
  summarise(afam_prop = prop(afam), 
            hispanic_prop = prop(hispanic), 
            other_prop = prop(other), 
            majority_prop = sum(prop(afam),prop(hispanic), prop(other))) 
  ## the proportions added up become more than 100%??
  
```

.

> Laying the groundwork, splitting up categories:

.

```{r}
fertility_modifiedB <- fertility_modified %>%
  group_by(morekids) %>% 
  summarise(
    afam_prop = mean(afam == "yes"),          
    hispanic_prop = mean(hispanic == "yes"),  
    other_prop = mean(other == "yes"),        
    
    ## since the proportions were added together and found to be more than one, identifying the overlapping categories
    
    afam_hispanic = mean(afam == "yes" & hispanic == "yes"),
    afam_other = mean(afam == "yes" & other == "yes"),
    hispanic_other = mean(other == "yes" & hispanic == "yes"),
    ##
    
    only_hispanic = hispanic_prop - sum(hispanic_other, afam_hispanic),
    
    only_afam = afam_prop - sum(afam_hispanic, afam_other),
    
    only_other = other_prop - sum(afam_other, hispanic_other),
    
    majority_prop = mean(afam == "no"& hispanic == "no"& other == "no"),
    minority_prop = sum(only_hispanic, only_afam, only_other, afam_hispanic, afam_other, hispanic_other))  

glimpse(fertility_modifiedB)
  

```

```{r}
## Attempting to make my own data set with simply the propotions calculated after summary
fertilitytwo <- readxl::read_excel("../../Data/fertilitytwo.xlsx")
fertilitytwo
## Quickly realising that won't work ;-;
```

.

> Identifying the proportion of the minority, where the majority is defined by those who do not identify as Hispanic, African American or Other.

.

```{r}
## majority vs minority
fertility_modifiedB %>%
  select(
    morekids, majority_prop, minority_prop
  ) %>%
  pivot_longer(cols = c(majority_prop, minority_prop),
               cols_vary = "fastest",
               names_to = "ethnicity") %>%
  mutate(
          ethnicity = factor(ethnicity,
                              levels = c( "majority_prop", "minority_prop"))
               ) %>% 
  gf_col(value ~ morekids,
         fill = ~ethnicity,
         position = "fill") %>%
  gf_labs(
    title = "Plot 5.B: Ethnicity: majority and minority vs morekids",
    subtitle = "position = fill"
  )
```

.

```{r}
## trying pivot longer 
fertility_modifiedB %>%select(-afam_prop, -hispanic_prop,-other_prop, -majority_prop)%>%
  pivot_longer(cols = c(only_hispanic, only_afam, only_other, afam_hispanic, hispanic_other),
               cols_vary = "fastest",
               names_to = "ethnicity") %>% 
  mutate(
                 ethnicity = factor(ethnicity,
                                    levels = c( "only_afam", "afam_hispanic", "only_hispanic", "hispanic_other", "only_other", ordered = TRUE ))
               ) %>% 
  gf_col(value ~ morekids,
         fill = ~ethnicity,
         position = "dodge") %>%
  gf_labs(
    title = "Plot 5.C: minority ethnicity vs morekids",
    subtitle = "position = dodge"
  )

```

```{r}
## PROPORTION
fertility_modifiedB %>%select(-afam_prop, -hispanic_prop,-other_prop, -majority_prop)%>%
  pivot_longer(cols = c(only_hispanic, only_afam, only_other, afam_hispanic, hispanic_other),
               cols_vary = "fastest",
               names_to = "ethnicity") %>% mutate(
                 ethnicity = factor(ethnicity,
                                    levels = c( "only_afam", "afam_hispanic", "only_hispanic", "hispanic_other", "only_other", ordered = TRUE ))
               ) %>%
   
  gf_col(value ~ morekids,
         fill = ~ethnicity,
         position = "fill")%>%
  gf_labs(
    title = "Plot 5.D: minority ethnicity vs morekids",
    subtitle = "position = fill"
  )
```

.

> -   **Understandings from graphs:**
>
>     -   Plot 5.B: Minority ethnicities have a higher proportion of those who want more kids.
>
>     -   Plot 5.C: There is no strong pattern to be concerned with in this plot, as some categories bleed into one another.
>
>     -   Plot 5.D: Those who identify as Hispanic make up for most of those in the minority ethnicity category who want to have more kids, even if by a small margin. The 'African American Only' and 'Other Only' categories are ever so slightly smaller than their proportion in the 'no' more kids category.

.

.

## Conclusion:
