---
title: "A2: Case Study 1"
author: "Vasantha M"
---

# Movie Profits

[Check out the Assignment here!](https://av-quarto.netlify.app/content/courses/analytics/casestudies/modules/20-movieprofits/)

.

## 1. Introduction

The data set observed in this case study goes over 3310 films, discussing their domestic and worldwide gross, production budget, genre, distributor and year of release, etc.

.

> ### 1.1. Overview
>
> TBA

### 1.2 Setup Chunk:

```{r}
library(tidyverse)
library(mosaic)
library(skimr)
library(ggformula)
library(kableExtra)
library(DT)
```

.

.

## 2. Examining Data

```{r}
mp <- read_delim(file = "../../Data/movie_profit.csv", delim = ";")
mp %>% glimpse()
```

```{r}
mp%>%inspect()
```

```{r}
mp %>% skim()
```

.

### 2. 1. Data Dictionary

There are `3310` rows and `10` columns in this data set.

.

#### 2.1.a. Qualitative Variables

-   `release_date` : When the movie was released. This data set contains movies released from 1936 through 2017.

-   `movie` : Tile of the film. This data set contains 3310 films.

-   `distributor` : Name of the company that distributed the film - there are six categories under this variable: Universal, Paramount Pictures, Warner Bros, and 20th Century Fox, Sony Pictures and Other. There are 42 missing entries.

-   `mpaa_rating` : The age-rating of each movie. There are four categories under this variable - R, PG-13, PG and G. This coloumn also has 130 missing entries.

-   `genre` : Describes the genre of the film, categorized into - Action, Adventure, Comedy, Drama, and Horror.

-   `decade` : Describes the decade in which the movie was released, i.e 1990, 2000, 2010 and so on. Though it is considered a quant type variable in the data set, it is technically qual variable.

.

#### 2.1.b Quantitative Variables

-   `production_budget` : The budget for the film in USD.

-   `domestic_gross` : The amount of money, in USD, that the film made across the US.

-   `worldwide_gross` : The amount of money, in USD, the film made across the world.

-   `profit_ratio` : This variable would technically be the (`worldwide_gross` - `production_budget`)/(`production_budget`), but the values in this section are 17 or so digits in nature. Could this be a misplaced decimal error?

    A simple calculation using the above formula on any film, suggests that the every value under profit_ratio must be multiplied by 10\^(-14).

.

### 2.2. Munging Data

```{r}
mp_mod <- mp %>%
  mutate(
    decade = as_factor(decade),
    movie = as_factor(movie),
    distributor = as_factor(distributor),
    distributor = replace(distributor, distributor == "", NA),
   
    mpaa_rating = as_factor(mpaa_rating),
    mpaa_rating = replace(mpaa_rating, mpaa_rating == "", NA),
    genre = as_factor(genre),
    
    profit_ratio = profit_ratio * 0.00000000000001
  )
glimpse(mp_mod)
```

.

```{r}
## viewing the dataset:
datatable(mp_mod, options = list(pageLength = 10))
```

.

### 2.3 Observations

.

#### Playing around with arrangement of certain coloumns

.

```{r}
## top 5 production budget:
mp_mod %>% arrange(desc(production_budget)) %>%
  head(5) %>%
  kbl(
   
    col.names = c(
      "Release\n Date", "Movie\n Name", "Production\n Budget", "Domestic\n Gross", "Worldwide\n Gross", "Distributor", "Rating", "Genre", "Profit\n Ratio", "Decade"
    ),
    caption = "Production Budget - Top 5"
  )%>%
  kable_styling(
    bootstrap_options = c(
      "striped", "hover",
      "condensed", "responsive"
    ),
    full_width = F, position = "center"
  )
```

.

```{r}
## top 5 domestic gross:
mp_mod %>% arrange(desc(domestic_gross)) %>%
  head(5) %>%
  kbl(
   
    col.names = c(
      "Release\n Date", "Movie\n Name", "Production\n Budget", "Domestic\n Gross", "Worldwide\n Gross", "Distributor", "Rating", "Genre", "Profit\n Ratio", "Decade"
    ),
    caption = "Domestic Gross - Top 5"
  )%>%
  kable_styling(
    bootstrap_options = c(
      "striped", "hover",
      "condensed", "responsive"
    ),
    full_width = F, position = "center"
  )
```

.

```{r}
## top 5 worldwide gross:
mp_mod %>% arrange(desc(worldwide_gross)) %>%
  head(5) %>%
  kbl(
   
    col.names = c(
      "Release\n Date", "Movie\n Name", "Production\n Budget", "Domestic\n Gross", "Worldwide\n Gross", "Distributor", "Rating", "Genre", "Profit\n Ratio", "Decade"
    ),
    caption = "Worldwide Gross - Top 5"
  )%>%
  kable_styling(
    bootstrap_options = c(
      "striped", "hover",
      "condensed", "responsive"
    ),
    full_width = F, position = "center"
  )
```

.

```{r}
## top 5 profit ratio:
mp_mod %>% arrange(desc(profit_ratio)) %>%
  head(5) %>%
  kbl(
   
    col.names = c(
      "Release\n Date", "Movie\n Name", "Production\n Budget", "Domestic\n Gross", "Worldwide\n Gross", "Distributor", "Rating", "Genre", "Profit\n Ratio", "Decade"
    ),
    caption = "Profit Ratio - Top 5"
  )%>%
  kable_styling(
    bootstrap_options = c(
      "striped", "hover",
      "condensed", "responsive"
    ),
    full_width = F, position = "center"
  )
```

.

> -   Its very interesting to note that the top 5 most expensive movies to make, The top 5 movies that made the most money worldwide and the top 5 movies that made the most in profits have not a single entry in common.
>
> -   Adventure and action films seem to make more money, as well as tend to cost more to make.
>
> -   horror and drama films seem to have the best profit turnout.

.

.

## 3. Plotting Data

### Q: Distributor vs Profit Ratio by Genre - How do profit ratios vary in terms of genre and distributor?

```{r}
mp_mod %>%
  group_by(distributor, genre) %>%
  summarise(median_profit_ratio = median(profit_ratio)) %>% 
  
  gf_col( genre  ~ median_profit_ratio ) %>% 
  gf_facet_wrap(~distributor)%>%
  gf_labs(
    title = "Profits made by Film Distributors",
    subtitle = "Ratio of Profits to Budgets",
    caption = "October 2024 "
  )
```

.

.

### Other Plots:

```{r}
## genre alone by profit ratio
mp_mod %>%
  gf_col(genre ~ profit_ratio)
```

.

```{r}
## genre alone by median_profit ratio
mp_mod %>%
  group_by(genre) %>%
  summarise(median_profit_ratio = median(profit_ratio)) %>% 
  gf_col(genre ~ median_profit_ratio)
```

.

```{r}
## distributor alone by profit ratio
mp_mod %>%
 gf_col(distributor ~ profit_ratio)
```

.

```{r}
## distributor alone by median_profit ratio
mp_mod %>%
  group_by(distributor) %>%
  summarise(median_profit_ratio = median(profit_ratio)) %>% 
  gf_col(distributor ~ median_profit_ratio)
```

.

> -   Why use median?
>
>     .
>
>     Because, of its effectiveness in representing the central tendency of skewed or non-normally distributed data.

.

.

## 4. Task and Discussion

.

### From the Course Website

.

-   ***Identify the type of plot***

    The above is a bar graph.

-   ***What are the variables used to plot this graph?***

    `distributor`, `genre`, `profit_ratio`

-   ***If you were to invest in movie production ventures, which are the two best genres that you might decide to invest in?***

    ```{r}
    ## genre alone by median_profit ratio
    mp_mod %>%
      group_by(genre) %>%
      summarise(median_profit_ratio = median(profit_ratio)) %>% 
      gf_col(genre ~ median_profit_ratio)
    ```

    Horror and Drama

-   ***Which R command might have been used to obtain the separate plots for each distributor?***

    `ggformula::gf_facet_wrap`

-   ***If the original dataset had BUDGETS and PROFITS in separate columns, what preprocessing might have been done to achieve this plot?***

    `mp_mod <- mp %>% mutate( ...., profit_ratio = PROFITS/BUDGETS)`

.

.

### My Journey with this Case Study:

TBA
