---
title: "(1) Day 1: Hello Universe"
author: "Vasantha M"
---

## Introduction

hello universe, hello world :D

> -   Creating Posts:
>
>     -   Navigate to "Data-As-Material-Class-Folder\>posts" and create a new folder
>
>     -   lets name the folder "Day X"
>
>     -   Then proceed to "File\>New File\>Quarto Document"
>
>     -   Title and Author your document
>
>     -   Save the new document IN "DAY X" as index.qmd EACH TIME.
>
>         -   There can be only 1 index file in each folder!

## My First Piece(s) of R Code

```{r}
#| label: setup
library(tidyverse)
library (ggformula)
library(babynames)
```

### Lets look at every baby name from the US starting all the way from...

### .....1880?!?!?

```{r}
babynames
```

### How many "Aditi"s have existed?

```{r}
babynames %>% filter(name =="Aditi") %>% 
  gf_line(n~year)

```

### What about "Adhithi"?

```{r}
babynames %>% filter(name == "Adhithi") %>% 
  gf_line(n~year)
```

### "Adithi"?

```{r}
babynames %>% filter(name == "Adithi") %>% 
  gf_line(n~year)
```

### or "Adhiti"?

```{r}
babynames %>% filter(name == "Adhiti") %>% 
  gf_line(n~year)
```

### In compiling them all into a single graph....

```{r}
babynames %>% filter(name == "Aditi"|name == "Adhithi"|name == "Adithi"|name == "Adhiti") %>% 
  gf_line(n~year)
```

### WOAHH!!

### Although, my own name has some dissapointing results :(

```{r}
babynames %>% filter(name == "Vasantha") %>% 
  gf_line(n~year)
```

### My mother's name seems to be faring better than mine, across the world..

```{r}
babynames %>% filter(name == "Meena") %>% 
  gf_line(n~year)
```

### As are her counterparts!!

```{r}
babynames %>% filter(name == "Mina") %>% 
  gf_line(n~year)
```

### And together...

```{r}
babynames %>% filter(name == "Meena"|name == "Mina") %>% 
  gf_line(n~year)
```

### :)
